---
title: "Course Project 1"
author: "Federico Vera"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1)Loading and preprocesing the data

## 1_1 Load the Data

We take the URL of the data and put it on a local directory with  a copy on Github, load the data into R



```{r url}
# URL address of the data
URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

# download the file 
download.file(URL,destfile="./datastore/repdata_2Fdata_2Factivity.zip",method="auto")


# Unzip the file
unzip(zipfile="./datastore/repdata_2Fdata_2Factivity.zip",exdir="data")
 
```
## 1_2 Process/Transform the data into a format suitable for analysis

Cleaning the NA data

```{r data}

data <- read.csv("data/activity.csv", header=TRUE)
clear_Data <- na.omit(data)

```


## 2) What is mean total number os steps taken per day?

## 2_1 Calculating Total number of steps per day

```{r steps}

steps_x_day <- aggregate(clear_Data$steps, by= list(Steps.Date= clear_Data$date), FUN = "sum")

```

## 2_2 Histogram of the total number os steps taken each day

plot the number of steps  in the x axis, divided in 20 breaks, and the y shows
the frecuency of how many days that quantity appear.

```{r  echo="FALSE" }
hist(steps_x_day$x,
     breaks=20,
     main="Histogram total number steps x day",
     xlab= "Steps x day",
     ylab= "n° of days with this steps"
     )
```
## Calculating and reporting the mean and median of the total number of 
##steps taken per day

```{R mean1}

#Mean
mean(steps_x_day$x, na.rm=TRUE)

```


```{R median1}
#Median
median(steps_x_day$x, na.rm=TRUE)

```
## 3) What is the average daily activity pattern?
## 3-1 Make a time series plot of the 5 minute interval vs average number of steps taken


```{R Plot daily activity pattern}
avg_steps_x_interval <- aggregate(steps ~ interval, clear_Data, mean)

plot(avg_steps_x_interval$interval, avg_steps_x_interval$steps, type="l",
     col=1,
     main="Average number of steps by Interval",
     xlab="time interval",
     ylab="Average number of steps")


```


## 3_2  Which 5-minute interval, contains the maximum number of steps?

```{R}
max_interval <- which.max(avg_steps_x_interval$steps)

print (paste("The interval with the highest avg step is",avg_steps_x_interval[max_interval,]$interval," and the interval's step number is",round(avg_steps_x_interval[max_interval,]$steps, digits=1)))

```
## 4) Imputing missing values

## 4_1 Calculate and report the total number os missing values

```{R missing_values}
missing_value_act <- data[!complete.cases(data),]
nrow(missing_value_act)
```

## 4_2 Devise a strategy for filling missing values in the dataset

# Im' going to replace the missing values (NA) with the average 5 minute interval

```{R }
data$full_steps <- ifelse(is.na(data$steps), round(avg_steps_x_interval$steps[match(data$interval,avg_steps_x_interval$interval)
],0),data$steps)

head(data)

```
 ## 4_3 Create a new dataset with the NA values filled
```{R}
# new dataframe

full_data <- data.frame(steps=data$full_steps, interval= data$interval, date= data$date)

head(full_data)
```
## 4_4 Make a Histogram of the total number of steps taken each day


```{r full steps}

full_steps_x_day <- aggregate(full_data$steps, by= list(Steps.Date= full_data$date), FUN = "sum")

```




plot the number of steps  in the x axis, divided in 20 breaks, and the y shows
the frecuency of how many days that quantity appear.

```{r  }
hist(full_steps_x_day$x,
     breaks=20,
     main="Histogram total number steps x day",
     xlab= "Steps x day",
     ylab= "n° of days with this steps"
     )
```

Calculating the mean and median


```{R mean2}

#Mean
mean(full_steps_x_day$x, na.rm=TRUE)

```


```{R median2}
#Median
median(full_steps_x_day$x, na.rm=TRUE)

```
The mean was reduced a litle bit, but the median keep the same value.
## Are there differences in activity patterns between weekdays and weekends?

## 5_1 Create a new factor variable in the dataset with two levels- "weekday" "weekend" indicating whether a given date is a weekday pr weekend day

```{r }

full_data$newDate <- as.Date(full_data$date, format = "%Y-%m-%d")

full_data$weekday <- weekdays(full_data$newDate)

full_data$day_type <- ifelse(full_data$weekday=='Saturday' | 
full_data$weekday=='Sunday', 'weekend','weekday')

head(full_data, n=12)
```

## 5_2 Make a panel plot containing a time series plot  of the 5-minute interval  and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r }

library(ggplot2)

avg_full_data <- aggregate (steps ~ interval + day_type, data = full_data, mean)

ggplot(avg_full_data, aes(interval, steps))+
        geom_line()+
        facet_grid(day_type ~.)+
        xlab("5-minute interval")+
        ylab("average n° steps")


```


